Who Stays Single? Exploring the Factors Behind Non-Marriage in the United States

### 2024-11-26 paper/paper.qmd
update reference

This paper examines the determinants of non-marriage by analyzing a rich dataset from IPUMS USA [@ipumsusa], which offers harmonized census and survey data spanning demographic, social, and economic variables. Prior studies have largely focused on individual aspects of marital behavior, such as income or education, but few have comprehensively analyzed the intersection of multiple socio-demographic factors. This study bridges that gap by exploring the combined effects of education, income, race, gender, and age on marital outcomes.

@@

The dataset was analyzed using R [@citeR] and utilized several R packages for data manipulation and visualization, including ggplot2 [@citeGGplot2], dplyr [@citeDplyr], and here [@citeHere]. The data was processed using the Apache Arrow package [@citeArrow] for efficient file handling and visualized with scales [@citeScales] for enhanced graphics. Reproducibility was ensured using knitr [@citeKnitr]. Data was extracted from IPUMS USA [@ipumsusa]. Guidance on storytelling with data was drawn from Telling Stories with Data [@tellingstories].

>>>>

This paper examines the determinants of non-marriage by analyzing a rich dataset from IPUMS USA [@ipumsusa], which offers harmonized census and survey data spanning demographic, social, and economic variables. Prior studies have largely focused on individual aspects of marital behavior, such as income or education, but few have comprehensively analyzed the intersection of multiple socio-demographic factors. This study bridges that gap by exploring the combined effects of socio-demographic factors. Specifically, the estimand of this study is the reason of an individual remaining unmarried, conditioned on their socio-demographic attributes, including education level, income, race, gender, and age. This focus allows for a nuanced understanding of how these factors interact to influence marital decisions.

@@

The dataset was analyzed using R [@citeR] and utilized several R packages for data manipulation, visualization, and analysis, including ggplot2 [@ggplot2] for creating elegant graphics, dplyr [@citeDplyr] for data wrangling, and here [@citeHere] for simplifying file management. The data was processed efficiently using Apache Arrow [@citeArrow], and visualizations were further enhanced with the scales package [@citeScales]. Bayesian modeling was implemented using rstanarm [@citeRstanarm], while statistical results were tidied with broom [@citeBroom] and broom.mixed [@citeBroomMixed]. Tables were constructed using kableExtra [@citeKableExtra] for a polished presentation. Reproducibility was ensured with knitr [@citeKnitr]. Data was extracted from IPUMS USA [@ipumsusa], and guidance on storytelling with data was drawn from Telling Stories with Data [@tellingstories].
